---
title: "Network3_ERGM"
author: "Jingyi Sun"
date: "10/2/2021"
output: html_document
---

We will see how to use ERGM models to do SNA.

The goal of ERGMs is to “describe parsimoniously the local selection forces that shape the global structure of a network” (Hunter et al. 2008).

"Ergm has the capability of approximating a maximum likelihood estimator for an ERGM given a network data set; simulating new network data sets from a fitted ERGM using Markov chain Monte Carlo; and assessing how well a fitted ERGM does at capturing characteristics of a particular network data set." (Hunter et al. 2008).

Let's first import packages.
```{r}
#install.packages("network")
#install.packages("ergm")
library(ergm)
library(network)
```

Read our follower-followee edgelists and individual-level covariates.

```{r}
el=read.csv("follow_el.csv",header=F, stringsAsFactors = F)
head(el)

cov=read.csv("artists2.csv",stringsAsFactors = F)
cov=cov[,-1]
# make a variable of tenure
cov$tenure=2021-cov$since
head(cov)

# to reduce modeling time, I only focus on older users with more than 1 year of tenure
old=cov$author[cov$tenure>0]
el=el[(el$V1 %in% old)&(el$V2 %in% old),]

```

Then we can buid the network and add node-level attributes.

```{r}
# build the network
net=network(el,directed = T)
plot(net)

# get an ordered list of nodes from the network
ordered=network.vertex.names(net)

# subset the covariates and order in the same way as ordered nodelist from the network
net_cov=cov[cov$author %in% ordered,]
net_cov=net_cov[match(ordered,net_cov$author),]

# set node attributes
network::set.vertex.attribute(net, names(net_cov),net_cov)
table(network::get.vertex.attribute(net,attrname = "tenure"))
```

Let's model this network one step at a time.
For a list of ergm terms, check here: https://rdrr.io/cran/ergm/man/ergm-terms.html

Edges MUST be included. 

First we add in the mechanism of reciprocity. 

```{r}
fit=ergm(net~ edges 
              +mutual
              ,control=control.ergm(MCMC.samplesize=5000,MCMLE.maxit=500))
summary(fit)
```
```{r}
# Does this model have satisfactory goodness of fit?
mcmc.diagnostics(fit)
gof <- gof(fit,control=control.gof.ergm(nsim=200)) 
par(mfrow=c(3,2))
plot(gof)

# GOF is not good.
```



Second, we add the mechanism of preferential attachment.


```{r}
# ERGM degenerates VERY FREQUENTLY!

fit=ergm(net~ edges 
         +mutual
         +idegree(1)
         +gwidegree(0, fixed=TRUE)
         ,control=control.ergm(MCMC.samplesize=5000,MCMLE.maxit=500
                               ,MCMLE.density.guard.min = 100000,MCMLE.density.guard = exp(3),
                               MCMC.burnin=100000, MCMC.interval=10000))
summary(fit)
```
```{r}
mcmc.diagnostics(fit)
gof <- gof(fit,control=control.gof.ergm(nsim=200)) 
par(mfrow=c(3,2))
plot(gof)

# GOF is better
```
Let's add the homophily effect of tenure. 


```{r}

fit=ergm(net~ edges 
              +mutual
              +idegree(1)
              +gwidegree(0, fixed=TRUE)
              +nodematch("tenure")
              ,control=control.ergm(MCMC.samplesize=5000,MCMLE.maxit=500
                     ,MCMLE.density.guard.min = 100000,MCMLE.density.guard = exp(3),
                     MCMC.burnin=100000, MCMC.interval=10000))
summary(fit)
```


